---
title: "gfdata `get_all' Vignette"
author: "Philina English"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gfdata Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  # cache = TRUE,
  cols.print = 9,
  rows.print = 9,
  eval = TRUE,
  autodep = TRUE,
  fig.path = "get-all-figs/",
  cache.path = "get-all-cache/"
)
```

```{r, echo=FALSE, eval = TRUE, cache=FALSE}
.error <- tryCatch(get_ssids(), error = function(e) "Error")
.eval <- !identical(class(.error), "character")
.eval <- TRUE
```


# Why use a `get_all*` function?

The `get_all*()` functions have been designed to retrieve all fishery-independent data relevant to a particular species, or set of species. 
If `ssid = NULL` they will return all fishery-independent data in the database, including trap (sablefish), longline, jig, and historic trawl surveys.  
If a character string of major stat area codes is provided to the argument `major`, than all sets or samples from that area will be returned. 
If retrieving data for multiple species at once, this function will be a lot faster than the original ones because sql scripts are called once rather than repeatedly for each species. 

## Non-standard data 

The other `get_*()` survey functions focus on retrieving sets and samples that conform to current design-based standards and survey grids. 
This means that they do not retrieve sets and samples from cells that fall within Rockfish Conservation Areas, or that have otherwise been excluded from the latest version of the survey grid. 
<!-- Their default settings also exclude tows that are shorter than the standard length.  -->

They also were not built to retrieve data that differed at the skate level (like gear comparison studies), which this function will return only when gear variables (currently checking for differences in hook code and size) differ between skates. 


## Duplication of fishing events and specimens

One risk in using these functions is that some fishing events and specimen ids may be duplicated in the resulting data set. 
This occurs because some vessel trips conduct sampling for multiple survey series, and unless joining is based on grouping codes (which aren't used consistently for some surveys ) the only way to connect a fishing event to a survey series id (SSID) is through the vessel trip. 
This can result in events and specimens getting assigned to both surveys that were conducted on the same trip. 
The most common instances of this have custom corrections coded into an internal function `correct_ssids()` that is applied automatically by the `get_all*` functions. 
Other ways duplication occurs can be due to missing covariates (e.g., both event level and survey defaults are missing for `doorspread_m` on a couple sets for some trawl series), or when multiple vials of DNA were collected and `return_dna_info = TRUE`. 


# Examples

If you don't already have the package installed, see the general gfdata vignette for instructions.
Here we will load the package along with dplyr.

```{r, cache=FALSE, warning = FALSE, message = FALSE}
library(gfdata)
library(dplyr)
```

The available arguments are described in the help documentation:

```{r, eval = FALSE}
?get_all_survey_sets
?get_all_survey_samples
```

 
## What survey data is available for a species?

As an example, we might want to determine what survey set data exists for Bluntnose Sixgill Sharks (*Hexanchus griseus*) in our database. For now, we will leave the default settings that pull all surveys and all areas. 
It should be noted that some data from United States waters is in our data base, so if we are only interested in Canadian records those sets would need to be filtered. 
It's not the default for this function, but we'll start by using `remove_false_zeros = FALSE` because the values returned will then match `get_survey_sets()` for any events returned by both functions. 
For any survey that encounters this species, this function will return all sets that did not encounter them as well. 
So, to make positive sets visible to `glimpse()`, We will sort by descending `catch_count` and `catch_weight`. 

```{r, eval=TRUE, warning=TRUE}
d0 <- get_all_survey_sets("Bluntnose Sixgill Shark", remove_false_zeros = FALSE)
d0 <- d0 |> arrange(-catch_count, -catch_weight)
head(d0)
```

Notice that `catch_weight` sometimes contains zeros when `catch_count` is at least 1. This may be because some specimens may be too large for the scale and null values are assumed by our SQL calls to be zeros. To avoid this, use the argument `remove_false_zeros = TRUE`. 

```{r, eval=TRUE, warning=TRUE}
d <- get_all_survey_sets("Bluntnose Sixgill Shark", remove_false_zeros = TRUE)
d <- d |> arrange(-catch_count, -catch_weight)
head(d)
```

So, which surveys encountered the most of this species? Some surveys only count individuals and others only weight the total catch, so we will summarize both count and weight variables.

```{r, message=FALSE}
d |> group_by(survey_series_id, survey_series_desc) |> 
  summarise(individuals = sum(catch_count, na.rm = TRUE),
            weight = sum(catch_weight),
            events = length(unique(fishing_event_id)),
            rows = n()) |>
  arrange(-individuals, -weight)
```

We can also tally the number of unique fishing events versus the number of rows of data returned. 
Data from surveys with more rows than events has been returned at the skate level. 
This will happen any time one of the surveys being returned has skates that differ in the gear (currently just working of differences in hook type) used within the same fishing event. 
This was the case for SSID 48, but this also results in all other surveys with skate level records returning data at this level too. 
If one preferred the other surveys with skate level data, which does not have hook differences between skates, to be summarized at the event level, thay could be called separately using the argument `ssid = c(14, 35)`.  

The vast majority of records come from the IPHC, followed by the Dogfish and Hard Bottom Longline surveys, both conducted in the Strait of Georgia (aka. Inside South). Because only the IPHC covers a wider area, we can explore the spatial distribution of catches within that survey only, and confirm that they are most frequently caught in the Strait of Georgia, major stat area "01". 

```{r, message=FALSE}
d |> filter(survey_series_id == 14) |> 
  group_by(major_stat_area_code) |> 
  summarise(individuals = sum(catch_count, na.rm = TRUE),
            weight = sum(catch_weight),
            events = length(unique(fishing_event_id))) |>
  arrange(-individuals, -weight)
```


## What survey samples are available for a species within a specific area?

As an example, we might want to determine what survey sample data exists for Pacific Spiny Dogfish in the Strait of Georgia. 
The area argument is `major` and accepts character vectors of major stat area codes. 
A table of options can be retrieved with `get_major_area()`. 

```{r, eval=.eval}
d2 <- get_all_survey_samples("north pacific spiny dogfish", major = c("01"))
```

```{r, eval=.eval, message=FALSE}
d2 |> group_by(survey_series_id, survey_series_desc) |> 
  summarise(specimens = length(unique(specimen_id)),
            lengths = sum(!is.na(length)),
            weights = sum(!is.na(weight)),
            age_structures = sum(age_specimen_collected)
            ) |>
  arrange(-specimens)
```

This should return all fishery-independent specimen records.
If you want to focus on specimens that come from design-based survey sets you can add arguments that filter for unsorted random samples that come from events that have grouping codes that match those expected for the current survey design. 
This can also be achieved by filtering for specimens where `!is.na(grouping_code)` and checking that the `sample_type_comment` and `sample_source_desc` notes are consistent with the specimens being from random samples.
Note: Some surveys do not use grouping codes, and therefore won't be returned when the `grouping_only` option is used. 
In this case, all of SSID 76 and 48 are missing. 

```{r, eval=.eval, message=FALSE}
d3 <- get_all_survey_samples("north pacific spiny dogfish", major = c("01"), 
                             unsorted_only = TRUE,
                             random_only = TRUE,
                             grouping_only = TRUE)

d3 |> group_by(survey_series_id, survey_series_desc) |> 
  summarise(specimens = length(unique(specimen_id)),
            lengths = sum(!is.na(length)),
            weights = sum(!is.na(weight)),
            age_structures = sum(age_specimen_collected)) |>
  arrange(-specimens)
```

If rather than limiting the specimens returned to design-based conventions, you want to retrieve covariates for use in model-based analyses change then set `include_event_info = TRUE`.
For example, when this is applied to the various longline surveys in the Strait of Georgia, one can test the effects of variables like depth, date, hook type and size on the sex and sizes of fish caught.

```{r, eval=.eval, message=FALSE}
d4 <- get_all_survey_samples("north pacific spiny dogfish", major = c("01"), ssid = c(39, 40, 48, 76),
                             include_event_info = TRUE)
```

```{r, eval=.eval, message=FALSE}
d4 |> group_by(survey_series_id, activity_desc, hook_desc, hooksize_desc) |> 
  summarise(specimens = length(unique(specimen_id)),
            lengths = sum(!is.na(length)),
            years = paste(min(year), "-", max(year))) |>
  arrange(-specimens)
```
These are the variables returned by default: 

```{r, eval=.eval, message=FALSE}
glimpse(d2)
```
And these additional variables were returned for longline surveys when `include_event_info = TRUE`:

```{r, eval=.eval, message=FALSE}
glimpse(d4[, !names(d4) %in% names(d2)]) 
```
A different set of variables would be returned if only retrieving samples from trawl surveys. 


## Surveys with overlapping stratifications

Some fishing events are assigned to multiple surveys, which may or may not be fully or partially overlapping, and are defined by the same activity code in the database. To get all sets with matching activity codes one can use `include_activity_matches = TRUE`. It will return all events that share the same `activity_code` as any SSIDs requested. 

```{r, eval=.eval, message=FALSE}
d5 <- get_all_survey_samples("north pacific spiny dogfish", ssid = c(7),
                             include_activity_matches = TRUE,
                             include_event_info = TRUE,
                             remove_duplicates = TRUE #> default
                             )

d5 |> group_by(survey_series_id, survey_abbrev, activity_desc) |> 
  summarise(specimens = length(unique(specimen_id)),
            years = paste(min(year), "-", max(year))) |>
  arrange(-specimens)
```

NOTE: This always occurs by default for the sablefish surveys (ssid = any of 35, 41, 42, 43), because SSIDs for that survey are confused by sharing trip ids. In order to separate the types of sablefish surveys one needs to split data from this survey by the `reason_desc` variable. 


