---
title: "Historical Rex Sole discards"
author: "EK"
date: "February 26, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(gfdata)
```

# Background
We want to estimate historical Rex Sole discards. Our approach is to apply a 
ratio of modern Rex Sole discards to modern landings of a commonly co-caught
species that was recorded well historically.

1. Pull catch for Rex and cocaught species (from 
gfdata::get_cocaught_species()).

```{r}
get_cocaught_species

rex <- get_cocaught_species(610)
head(rex)
```

Here it appears Dover Sole and Pacific Cod are most frequently caught with 
Rex Sole. I expect Pacific Cod to have better records of historical landings,
so will build modern Rex discard ratios relative to Pcod landings (as is done
in reconstructing rockfish historic discards). 


2. Sum annual catch of Rex and cocaught (here, Pcod) landings and discards (from
groundfish trawl fishery 1996-2006), and calculate annual ratio of target 
discard to cocaught landings.


```{r}

cocaught_ratios

rex_pcod <- cocaught_ratios(610, 222, start_year = 1996, end_year = 2006, 
  sector = "groundfish trawl")
head(rex_pcod)
```

3. Next, select trusted reported years, calculate overall mean ratio (mean of
annual means, as is done for rockfish), and apply to desired historic catch 
years.

# Some questions:

Would a different reference species be better for this purpose? 

Do we also want to calculate historical Rex landings in this way? 

Or, if we trust our historical Rex landing records, do we want to base our 
discard estimates on those instead of a co-caught species?

